---
title: "How to Lie with Statistics"
format: 
  dashboard:
    logo: "images/hex_logo.png"
    orientation: columns
    nav-buttons:
      - icon: github
        href: https://github.com/samhendersonpalmer
      - icon: linkedin
        href: https://www.linkedin.com/in/samhendersonpalmer/
---

# Built in bias

## Column {width = "65%"}

### Row {height = "10%"}

Selection bias e.g. asking a room full of gun owners if guns are bad for society

When talking about the 'average person', ask yourself "how could anyone have found that out?"
- The result of a sampling study no better than the sample it's based on
- Must be a representative sample

Test of a random sample
- Does every name or thing in the whole group have an equal chance to be included in the sample?

### Row {height = "70%"}

Placeholder text for column 1, row 2

### Row {height = "20%"}

Placeholder text for column 1, row 3

## Column {width = "35%"}

::: {.panel-tabset .nav_pills}

## Show me reality

```{r, loadpackages}
#| echo: false
#| include: false
library(tidyverse)
```

```{r, printstarwars}
library(waffle)

theme_set(theme_void())

gun_toters <- data.frame(gun_owner = rbinom(100, 1, 0.32))

gun_toters |> mutate(gun_owner = case_when(gun_owner == 1 ~ "Gun owner",
.default = "Not a gun owner")) -> gun_toters

gun_toters|> 
count(gun_owner) |> 
ggplot(mapping = aes(fill = gun_owner, values = n)) +
  geom_waffle(n_rows = 10,
    size = 1,
    color = "white",
    make_proportional = TRUE,) +
scale_fill_manual(values = c("#d61c1d", "#eceaea")) +
  labs(
    title = "Gun owners",
    x = NULL, y = NULL, fill = NULL) +
coord_equal() +
theme(legend.position = "bottom",
      plot.title = element_text(hjust = 0.5))

```

## Lie to me
```{r plottcars}
plot(cars)
```
:::
